---
title: "Exercise session 3"
author: "Simon"
date: "9/12/2021"
output: html_document
---


```{r}
library(pacman)
p_load(tidyverse,rio)
dat <- import("http://fmwww.bc.edu/ec-p/data/wooldridge/wage1.dta")
dat %>% glimpse()
```
# 1. Create two scatterplots, one for education and wage, and one for experience and wage. Does it, in this simple descriptive analysis, look like a relation exist between wage and each of the two variables?


```{r}
p_load(gridExtra)
```


```{r}
p1 <- dat %>%
ggplot(aes(x = educ, y = wage)) +
geom_point(color = "darkgreen",
alpha = 0.7) +
theme_classic()



p2 <- dat %>%
ggplot(aes(x = exper, y = wage)) +
geom_point(color = "darkgreen",
alpha = 0.7) +
theme_classic()
```

```{r}
grid.arrange(p1, p2, nrow = 1)
```
There appears to be a positive relation between education and wages. There is possibly one between experience and wages. Although, the latter is less apparent from a casual look at the data.



# 2. Obtain estimates of $\beta_0$, $\beta_1$ , and \beta_2 using OLS

```{r}
mod1 <- lm(wage ~ educ + exper, data = dat)
summary(mod1)
```

# 3. Interpret your coefficient estimates

• *Intercept*: We find an estimate of $-3.39. The estimated model suggest that were both education and
experiences to be zero, wages would be negative. The person so endowed would have to pay to be on
the labor market! Luckily, it rarely happens that an individual has neither education nor experience.

• *educ*: The estimated coefficient for education is 0.64. This suggest that when education increases by
one year, hourly wages increases on average by $0.64.

• *exper*: The estimated coefficient for experience is 0.07. The estimate suggests that when experience
increases by one year, hourly wages increases on average by $0.07.

# 4. Using your estimated regression equation, calculate the predicted values of your model for the values of educ and exper in the dataset.

```{r}
## Coefficients from model
coefs <- coef(mod1)
coefs
```


```{r}
dat <- dat %>%
mutate(y_hat_1 = coefs[1] + coefs[2] * educ + coefs[3] * exper)
```

```{r}
dat$y_hat_2 = predict(mod1)
```

```{r}
sum(dat$y_hat_1 != dat$y_hat_2)
```
There is no difference between the predicted values using either method.

# 5. Create a scatterplot with actual values of wage and predicted values of wage. Are there signs that the model fit better for some wage values than other?


```{r}
dat %>%
ggplot(aes(x = wage, y = y_hat_1)) +
geom_point(color = "darkgreen",
alpha = 0.7) +
geom_abline(slope = 1, intercept = 0) +
lims(x = c(0, 25),
y = c(0, 25)) +
theme_classic() +
labs(x = "Actual wage", "Predicted wage")
```




# 6. Show the equation of the residuals. Using your equation, calculate the residuals.

```{r}
dat <- dat %>%
mutate(res = wage - y_hat_1)
```

# 7. Show the distribution of the residual (histogram). Does it look like a normal distribution?


```{r}
hist(mod1$residuals, breaks=10, probability=T ,col="blue", xlab="res"); lines(density(mod1$residuals), col="red", lwd=2)
```



```{r}
dat %>%
ggplot(aes(x = res)) +
  geom_histogram(aes(y = ..density..), fill = "darkgreen", ) +
  stat_function(fun = dnorm, args = list(mean = mean(dat$res),sd = sd(dat$res)),color = "royalblue") +
  lims(x = c(-10, 15)) +
  theme_classic()

```


# 8. Test whether each of educ and exper has a statistically significant relation with wage in the estimated model.

We can observe the t-statistics directly in the output.

```{r}
summary(mod1)
```
We can also calculate them by hand instead

```{r}
t_educ <- ( 0.64427 - 0 ) / 0.05381
t_exper <- ( 0.07010 - 0 ) / 0.01098
c(t_educ, t_exper)
```
The critical values for the hypothesis tests

```{r}
n <- mod1$residuals %>% length()
k <- mod1$rank
df <- n - k - 1
t_lower <- qt(0.025, df = df)
t_upper <- qt(0.975, df = df)
c(t_lower, t_upper)
```

We can also calculate p-values

```{r}
p_educ <- 1 - pt(t_educ, df = df)
p_exper <- 1 - pt(t_exper, df = df)
c(p_educ, p_exper)
```
The p-values are both far below 0.05, the statistical significance level chosen for the test. As a result, we can
reject both null hypotheses.


# 9. Assume you are interested in testing the hypothesis whether the effect of educ on wage is equal to 0.5 or not.

The null hypothesis is
$H0 : \beta_1 = 0.5, H1 : \beta_1 \neq 0.5$
I will test the hypothesis at a 5 percent significance level.


Calculate t-values

```{r}
t_stat <- ( 0.64427 - 0.5 ) / 0.05381
c(t_stat, t_lower, t_upper)
```




