---
title: "Lecture 5"
author: "Simon"
date: "3/11/2021"
output: html_document
---

# Cumulative distribution function

```{r}
x=seq(1, 1000, 1); y=dunif(x, 0, 1000)


curve(dunif(x, 0, 1000), from = 0, to = 1000, bty="n", ylim=c(0.000, 0.0012))

```



```{r}
punif(750 , min = 0 , max = 1000) - punif(250 , min = 0 , max = 1000)
```



```{r}
curve(dunif(x, 250, 750), from = 0, to = 1000, bty="n", col= "blue")

```


# Probability density function


```{r}
x <- c(-2,seq(-2,2,0.01),2)
y <- c(0,dnorm(seq(-2,2,0.01)),0)
curve(dnorm(x,0,1), bty="n", xlim=c(-3,3)); polygon(x,y,col='skyblue')
```


```{r}
x=seq(1, 100, 1)
curve(dunif(x, 0, 100), from = 0, to = 100, bty="n", ylim=c(0, 0.012)) 
```


# Expectation (mean) for continous random variable

# Normal distribution

```{r}
curve(dnorm(x,0,1), bty="n", xlim=c(-3,3))
```


```{r}
x <- seq(-6,6,length=500)
plot(x,dnorm(x,mean=0,sd=1),type = "l",lty=1,lwd=3,col="blue"); curve(dnorm(x,0,1.5),add=TRUE,lty=2,col="blue", lwd=3)


```
Same mean different SD

```{r}
x <- seq(-6,6,length=500)
plot(x,dnorm(x,mean=0,sd=1),type = "l",lty=1,lwd=3,col="blue"); curve(dnorm(x,0.5,1),add=TRUE,lty=2,col="blue", lwd=3)

```

```{r}
library(mnormt)
mu <- c(0,0); sigma <- matrix(c(2,0,0,2),2,2)
x<-seq(-4,4,0.1); y<-seq(-4,4,0.1)
f<-function(x,y){dmnorm(cbind(x,y), mu, sigma)}; z<-outer(x,y,f)
persp(x,y,z, box=T, ticktype="detailed", theta=47, phi=6, expand = 0.65)

```


## Example 

A company produces lightbulbs whose life follows a normal distribution, with a
mean of 1,200 hours and a standard deviation of 250 hours. If we choose a
lightbulb at random, what is the probability that its lifetime will be between 900
and 1300 hours?

- First calculate the corresponding z-score of the upper limit (1300 hours)
and lowe limit (900 hours)

```{r}
z_low = (900 - 1200)/250; z_high= (1300 - 1200)/250

pnorm(z_high, mean = 0, sd=1) - pnorm(z_low, mean = 0, sd=1)

```

We could also do this since it is normaly distributed

```{r}
pnorm(1300, mean = 1200, sd=250) - pnorm(900, mean = 1200, sd=250)

```

Assume we are instead given a probability 0.1151, and we are asked to calculate
lifetime of bulbs a this probability


```{r}
qnorm(0.1151, mean = 1200, sd=250)

```

## Mini-Exercise

Life span of people in a Aalborg is normally distributed with a mean 82
years and standard deviation 6. What is the probability that a randomly
selected person from this class in Aalborg will live more than 90 years?


We can do it just using the normal distribution 

```{r}
1- pnorm(90, mean = 82, sd=6)
```
we can also transform to a standard normal distribution first

```{r}
z_score = (90 - 82)/6

1- pnorm(z_score, mean = 0, sd=1)
```

# Joint distribution


# Correlation

We can very easily calculate mean, variannce, covariance, correlation in R:

```{r}
x=rnorm(100, 0, 2)
y=rnorm(100, 1, 1)

var(x) # variance of x

mean(y) # mean of y

sd(x) # standard deviation of x

cov(x,y) # covariance between x and y

cor(x,y) # correlation between x and y

```

# Exxercise 

This is time for you to start dealing with real world data.
Download quarterly GDP and unemployment rate data for Denmark from
1995–2016.
I Data can be downloaded Statistics Denmark

https://www.statistikbanken.dk/statbank5a/selectvarval/saveselections.asp
https://www.statistikbanken.dk/statbank5a/selectvarval/define.asp?PLanguage=0&subword=tabsel&MainTable=NAHO2&PXSId=219819&tablestyle=&ST=SD&buttons=0


```{r}
library(readxl)
data <- read_excel("data lecture 5.xlsx")

y= data$BNP
u= data$U
```
**Calculate GDP growth. Plot the distribution of GDP growth and fit a probability density curve. Does it look like a normal distribution?**

```{r}
y_growth= diff(y)/y[-1]*100



h<-hist(y_growth, breaks=10, col="blue", xlab="BNP growth rate")
xfit<-seq(min(y_growth),max(y_growth),length=21)
yfit<-dnorm(xfit,mean=mean(y_growth),sd=sd(y_growth))
yfit <- yfit*diff(h$mids[1:2])*length(y_growth)
hist(y_growth, breaks=10, col="blue", xlab="BNP growth rate"); lines(xfit, yfit, col="red", lwd=2)
```
Hard to say with such a small amount of observations, but looks left skewed duo to the one observation of -5%. 


**Calculate the mean, variance and the standard deviation of both GDP growth and unemployment rate**

y_growth

```{r}
#mean
mean(y_growth)

# variance

var(y_growth)

#standard deviation

sd(y_growth)
```
Unemployment

```{r}
#mean
mean(u)

# variance

var(u)

#standard deviation

sd(u)
```


**Calculate the covariance and correlation of unemployment rate and GDP growth. Is the correlation positive or negative?**

```{r}
#We need to first remove the first observation from unemployment
length(u)
u= u[2:22]

#Covariance
cov(y_growth, u)
#Correlation
cor(y_growth, u)
```


**Show the relationship between GDP growth and unemployment in a scatter plot (with a line).**

```{r}
data=cbind(y_growth, u)

lm(y_growth~u)

plot(data, type = "p"); abline(lm(u~y_growth))
```

