---
title: "Lecture 8"
author: "Simon"
date: "18/11/2021"
output: html_document
---

# Student's t distribution

```{r}
qt(1-0.025, df = 10)
```
## Exercise 

```{r}

data = read.csv("ryder.csv", header = T)
x = as.numeric(data$hoejde) # we tell R this is a numerical data
x = na.omit(x) # We leave out individuals for whom data is missing
```


```{r}
n <- length(x) # calculates the no. of observations (n)
m <- mean(x) # calculates the sample mean
s <- sd(x) # calculates the sample standard deviation
se <- s/sqrt(n) # calculates the standard error
tscore <- qt(1-.025, df = n-1) # calculates the t score
ME <- tscore*se # calculates the margin of error (ME)
m + c(-ME,ME) # Calculates the lower and upper confidence limits of the mean of x
```

We can get this directly in R
```{r}
t.test(x)
```


# Confidence Intervals for Population Proportion

We use the same dataset but now for proportions

```{r}
y = table(data$koen)
y # we tell R this is categorical variable
```
we can get the proportions

```{r}
prob=y/sum(y)
```
Calculate using functions from slides

```{r}
p= as.numeric(prob[1]); p

y = data$koen # we tell R this is a numerical data
y = na.omit(y) # We leave out individuals for whom data is missing
n_y=length(y)

se= sqrt((p*(1-p))/n_y)
zscore <- qt(1-.025, df= (n-1))# ik sikker om det er z eller t
me= se*zscore
p + c(-me,me) # Calculates the lower and upper confidence limits of the mean of x
```

```{r}
prop.test(y, correct = FALSE)
```

# Condfidence Intervals for Population Proportion






# Confidence interval for variance

Assume we have a sample of 20 car dealers in Denmark, and we measure the
number of weekly car sales. The standard deviation of the car sales is 3.
Calculate the confidence interval of variance at 90%, 95%, and 99%?

We have all the values but we need to check the value (in blue) from
chi-square dist. table (here I will use R for only 95% confidence interval:)

First, I calculate the (critical) value for lower tail at 95% with 19
degrees of freedom (n − 1).

```{r}
qchisq(0.025, df=19)
qchisq(1-0.025, df=19)
```

We can now calculate the confidence interval 

Lower limit:

```{r}
n=20 # size of the sample
s2=9 # variance of the sample
(n-1)*s2/32.85
```
upper limit: 

```{r}
(n-1)*s2/8.907
```

# Sample size determination


# Exercise

**Discussion question: In Chi-square case, you noticed that we calculated the upper tail and lower tail values. Why did we not do that in the normal dist and student’s t distribution**

Because Normal and T dist. are symmetric about the mean where as chi-square is not. In other words the lower tail and upper tail for normal and t dist. will return the same value but opposite signs, e.g.,

```{r}
qnorm(0.025); qnorm(1-0.025)
```

```{r}
qt(0.025, df=30); qt(1-0.025, df=30)
```

```{r}
qchisq(0.025, df=30); qchisq(1-0.025, df=30)
```

**Assume a sample of size (n=29) with mean 20 and variance of the population is known to be 500? Calculate the 95% confidence interval of mean**

We can directly use the normal distribution case since population variance is known.

Confidence intervals can be calculated as follows:

1- calcculate standard errors: 

```{r}
n=29 # sample size
sigma2= 500; sigma=sqrt(500) # population variance and standard deviation
m= 20 # mean of the variable
```


```{r}
se=sigma/sqrt(n)
```


2. calcculate the z values

```{r}
z=qnorm(1-0.025) # value of z_a/2 at 95% confidence level
```


3. calculate the mean error

```{r}
me=se*z
```


4. calcualte the upper and lower limit

```{r}
m + me; m - me
```
Same ccould be done in one call

```{r}
m + z*sigma/sqrt(n); m - z*sigma/sqrt(n)
```


**Assume we do not have the population variance, and the sample variance is 450. Calculate the 95% confidence interval of mean.**

We cannot directly use the normal distribution case since population variance
is unknown. But we are given sample variance, so we can proceed to use
student’s t distribution in this case

We do the same just using a t dist instead of a normal dist. 

```{r}
t=qt(1-0.025, 28) # value of t_a/2 at 95% confidence level with 28 degrees of freedom
n=29 # sample size
s2= 450; s=sqrt(450) # population variance and standard deviation
m= 20 # mean of the variable
```


```{r}
m + t*s/sqrt(n); m - t*s/sqrt(n)
```


**Calculate the 99% confidence interval of the variance**

The formula for calculating confidence interval for variance is on slide 25


```{r}
chi_lower=qchisq(0.005, df=28) # value for lower tail dist. at 99%
chi_upper=qchisq(1-0.005, df=28) # value for upper tail dist. at 99%
```

```{r}
(n-1)*s2/chi_lower; (n-1)*s2/chi_upper
```


**Calculate the 90% confidence interval of the variance of height in the example**

```{r}
data = read.csv("ryder.csv", header = T)
x = as.numeric(data$hoejde) # we tell R this is a numerical data
x = na.omit(x) # We leave out individuals for whom data is missing
```

```{r}
n <- length(x) # calculates the no. of observations (n)
m <- mean(x) # calculates the sample mean
s2 <- var(x); s2 # calculates the sample variance
```
```{r}
chi_lower=qchisq(0.05, df=n-1) # value for lower tail dist. at 90%
chi_upper=qchisq(1-0.05, df=n-1) # value for upper tail dist. at 90%
```

```{r}
(n-1)*s2/chi_lower; (n-1)*s2/chi_upper
```


**Plot the distribution of heights. What kind of distribution does it look like?**

```{r}
hist(x, prob=T, col = rgb(0.8, 0.8, 1)); lines(density(x, adjust=2), lwd = 2, col="red")
```

