---
title: "Lecture 7"
author: "Simon"
date: "11/11/2021"
output: html_document
---

# Sampling distribution

```{r}
age=c(18, 19, 22, 21 ) # Mikael, Tanja, Tine, Simon
```

```{r}
samples <- combn(age, 2) # this calculates all possible combinations when we randomly select 2 samples
```

```{r}
samp_mean <- colMeans(samples)

x=table(samp_mean); x
```

```{r}
prop.table(x)
```
```{r}
barplot(prop.table(x), col = rgb(0.8, 0.8, 1))
```
## Another example 

```{r}
set.seed(312); y=sample(1:30, 20)
```

```{r}
samp_1=combn(y, 4); samp_1_m <- colMeans(samp_1) # sample with 4 obs.
samp_2=combn(y, 10); samp_2_m <- colMeans(samp_2) # Sample with 10 obs.
```

```{r}
hist(samp_1_m, col = rgb(0.8, 0.8, 1), xlim=c(2,28), xaxt='n', main="case 1 (random samples (n=4))", prob=T); axis(side = 1, at = seq(2, 28, by = 2)); lines(density(samp_1_m), lwd=2, col="red")
hist( samp_2_m, col = rgb(0.8, 0.8, 1), xlim=c(2,28), xaxt='n', main="case 2 (random samples (n=10))", prob=T); axis(side = 1, at = seq(2, 28, by = 2)); lines(density(samp_2_m), lwd=2, col="red")
```

```{r}
mean(y); mean(samp_1_m); mean(samp_2_m)
```

# Central limit theorem

```{r}
set.seed(213)
pop_1 <- rnorm(1000, mean=4.9, sd = 2.1); pop_2 <- rf(1000, 10, 20)
pop_3 <- runif(10000000, 1, 1000);
```


We can see it is 3 differnet distributions
```{r}
hist(pop_1, col="red", main = "Normal distribution")
hist(pop_2, col="green", main = "Right-skewed")
hist(pop_3, breaks=10, col="orange", main = "Uniform distribution")
```

We can take 1000 random samples with 5 obs. 
```{r}
x_samp5 = replicate(1000, mean(sample(pop_1, 5, replace = T)))
y_samp5 = replicate(1000, mean(sample(pop_2, 5, replace = T)))
z_samp5 = replicate(1000, mean(sample(pop_3, 5, replace = T)))
```


```{r}
hist(x_samp5, prob=T, col="red", main="1000 samples from normal dist. (n=5)"); lines(density(x_samp5))
hist(y_samp5, prob=T, col="green", main="1000 samples from right-skew dist. (n=5)"); lines(density(y_samp5))
hist(z_samp5, prob=T, col="orange", main="1000 samples from uniform dist. (n=5)"); lines(density(z_samp5))
```
We can now do the same with 15 obs. 

```{r}
x_samp15 = replicate(1000, mean(sample(pop_1, 15, replace = T)))
y_samp15 = replicate(1000, mean(sample(pop_2, 15, replace = T)))
z_samp15 = replicate(1000, mean(sample(pop_3, 15, replace = T)))
```


```{r}
hist(x_samp15, prob=T, col="red", main="1000 samples from normal dist. (n=15)"); lines(density(x_samp15))
hist(y_samp15, prob=T, col="green", main="1000 samples from right-skew dist. (n=15)"); lines(density(y_samp15))
hist(z_samp15, prob=T, col="orange", main="1000 samples from uniform dist. (n=15)"); lines(density(z_samp15))
```


# Sampling distribution of sample proportion



# Sampling dist. of sample variance

## Chi-Square distribution

Population mean 50 and population variance 25 ($\sigma$ = 5) The sample size is 10 (we draw a new random sample 5000 times)

```{r}
pop <- rnorm (100000 , mean = 50 , sd = 5 )
variances <- replicate (5000 , var (sample (pop, 10 )))
hist (variances, prob=TRUE, col="skyblue" )
```

Let us look at the distribution using eq on slide 29

```{r}
hist(variances*(10-1)/var(pop), prob=TRUE, col="skyblue"); curve(dchisq(x, df = 9), from = 0, to = 35, add = TRUE)
```
## Exxample

```{r}
1 - pchisq(14.4, df = 9)
```

```{r}
qchisq(0.05, df = 10); qchisq(0.95, df = 10)
```


# Exercise


**Create a population, which is normally distributed with the mean and standard deviation of your choice (as I have done on slide no. 28)**

```{r}
set.seed(213)
pop <- rnorm(1e+05, mean = 5, sd = 5)
```


**Draw a new random sample 1000 times (the sample size should be 30, i.e., n=30) and calculate sample variances**

```{r}
sam_var <- replicate(1000, var(sample(pop, 30)))
```


**Prove that E(s2) = 2 and V ar(s2) = V ar(s2) = 24 n−1 where s2 refers to the sample variances, and 2 refers to the population variance**

- E(s^2)
```{r}
mean(sam_var) 
var(pop)
```

- Var(s^2)
```{r}
var(sam_var)
```
```{r}
2 * (var(pop)^2)/29
```


**Plot the dist. of s2(n − 1)/2. Does it look like a chi-square distribution?**

```{r}
hist(sam_var * (30 - 1)/var(pop), breaks = 20, prob = TRUE, col = "skyblue"); curve(dchisq(x, df = 29), from = 0, to = 55, add = TRUE)
```



**Calculate the degrees of freedom in this example**

```{r}
30-1
```

**For the degrees of freedom that you have calculated, what is the chi-square value at the lower 0.01 interval and the upper 0.05 interval.**

```{r}
qchisq(0.01, df = 29); qchisq(0.95, df = 29)
```



**What is the probability that a sample (out of 1000) will have a variance below 35?**

We need to calculate the equation 

$$s^2 (n-1)/\sigma^2$$
using $s^2 = 35$

```{r}
35*(30-1)/var(pop)
```
Now we have the chi-score value and can calculate the change

```{r}
pchisq(40.91, df = 29)
```



**What is the probability that a sample (out of 1000) will have a variance above 15?**

We do the same with the new condition


```{r}
15*(30-1)/var(pop)
```
Now we have the chi-score value and can calculate the change, as we want the area to the left we take 1- the chance

```{r}
1- pchisq(17.53, df = 29)
```

**Repeat question no. 2 but this time draw a new random sample 1000 times (the sample size should be 5, i.e., n=5).**

```{r}
sam_var1 <- replicate(1000, var(sample(pop, 5)))
```


**Plot the dist. of s2(n − 1)/2. Does it look like a chi-square distribution? Comment on how the distribution has changed when we decreased the degrees of freedom.**

```{r}
curve(dchisq(x, df = 4), from = 0, to = 55, add = TRUE)
```


