---
title: "Lecture 4"
author: "Simon"
date: "7/10/2021"
output: html_document
---


# Joint probability distribution 

```{r}
volley <- matrix(c(30, 20, 60, 30, 20, 20), ncol = 2, byrow = TRUE)
colnames(volley) <- c("Male", "Female")
rownames(volley) <- c("Current", "Former", "Never")
volley1 <- as.table(volley)
prop.table(volley)

```


# Marginal probability distribution 

```{r}
addmargins(prop.table(volley))
```

# Conditional probability distribution 

```{r}
prop.table(volley, margin = 2) # Columns are given 

```
```{r}
prop.table(volley, margin = 1) #Rows are given
```

# Mean, variance and covariance of joint probability dist. 

Create data table
```{r}
data <- matrix(c(0.17, 0.11,
                 0.33, 0.17,
                 0.11, 0.11),ncol = 2, byrow = TRUE)
# Define column names
colnames(data) <- c("1", "2")
# Define row names
rownames(data) <- c("1", "2", "3")
data <- as.table(data)
data
```
Calcualte merginal probabilities 

```{r}
table <- addmargins(prop.table(data))
table
```
Create vectors that you need

```{r}
y=c(1,2,3)

py2= data[,2]

```
calculate mean

```{r}
mean_y= sum(y*py2/0.39)
mean_y
```
Calculate variance 

```{r}
var_y=sum((y-mean_y)^2*(py2)/0.39)
var_y
```


**Statistical independence**


**Mean of variables with joint probability dist.**

```{r}
data <- matrix(c(0.112, 0.074, 0.224,
                 0.112, 0.074, 0.074,
                 0.104, 0.134, 0.086), ncol = 3, byrow = TRUE)
# Define column names
colnames(data) <- c("2", "4", "6")
# Define row names
rownames(data) <- c("1", "2", "3")
data <- as.table(data)
prop.table(data)

```
```{r}
table <- addmargins(prop.table(data))
table
```
If we want to calculate the mean of Y take the marginal probabilities multiply by discrete values. 

```{r}
y=c(1,2,3)

py=table[1:3,4]
```

```{r}
mean_y= sum(y*py)
mean_y
```
Calculate the variance

```{r}
sum((y-mean_y)^2*py)
```

# Exercise 


**Recreate this table in R**

```{r}
data <- matrix( c (0.066, 0.044, 0.132, 0.066,
                   0.044, 0.044, 0.033, 0.044,
                   0.121, 0.099, 0.084, 0.073,
                   0.026, 0.053, 0.042, 0.024), ncol = 4, byrow = TRUE)
# Define column names
colnames(data) <- c("1", "3", "5", "7")
# Define row names
rownames(data) <- c("1", "2", "3", "4")
data <- as.table(data)
data<- prop.table(data)
data
```


**Calculate the marginal probabilities of X and Y**

```{r}
table <- addmargins(data)
table
```


**Calculate the expected value of X and Y , i.e., E[X] and E[Y ]**

Mean for X
```{r}
x=c(1,3,5,7)

px=table[5,1:4]

mean_x= sum(x*px); mean_x

```
Mean for Y

```{r}
y=c(1,2,3,4); py=table[1:4,5]
mean_y = sum(y*px); mean_y

```


**Given the condition that we have x = 3, what is the probability of getting y=2, i.e., Calculate P(y = 2|x = 3)?**

```{r}

joint_p= table[2,2]
marg_x3= table[5,2]
p_y2_x3= joint_p/marg_x3; p_y2_x3
```
Or use the table: 

```{r}
tb=prop.table(data, margin=2)

tb[2,2]
```


**Calculate the conditional mean and variance of X when y = 2?**

```{r}
x=c(1,3,5,7)

px2=table[2,1:4]

marg_y2= table[2,5]

#Mean
x_cond_mean= sum(x*px2/marg_y2);x_cond_mean

#Variance
sum((x - x_cond_mean)^2*(px2)/marg_y2)

```

**Calculate the variance of X and Y**

We use the mean from before

```{r}
mean_x
mean_y
```
```{r}
var_x=sum((x - mean_x)^2*(px));var_x
var_y=sum((y - mean_y)^2*(py));var_y

```



**Calculate the covariance between X and Y ?**

```{r}
cov_xy = 1*(1*0.066 + 3*0.044 + 5*0.132 + 7*0.066) +
  2*(1*0.044 + 3*0.044 + 5*0.033 + 7*0.044) +
  3*(1*0.121 + 3*0.099 + 5*0.084 + 7*0.073) +
  4*(1*0.026 + 3*0.053 + 5*0.042 + 7*0.024) - mean_x*mean_y
cov_xy
```

