---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
library(plotrix)
library(qcc)
library(scatterplot3d)


```

#Installing R and Rstudio

#Rstudio Basics

```{r}
4.6 * (2 + 3)^4

############
a=4
a
############

b= c(2,5,7)
b
###########

aa= matrix(c(6,4,8,1), ncol = 4, nrow = 4)
aa
#############
aaa=matrix(c(6,4,8,1,
             1,2,3,4,
             3,4,2,4,
             5,2,5,7), ncol = 4, nrow = 4)

aaa

```

```{r}
x <- 1:7
x

y <- seq(1, 20, by = 2)
y

y[c(2, 3, 4)]

length(y)

```

#Statistics Basics

#Graphs to descripe Categorical variables

**Frequency Distribution**

```{r}
volley <- matrix(c(50,90,40, (50/180)*100, (90/180)*100, (40/180)*100), ncol = 2, nrow = 3)
colnames(volley) <- c("Membership", "Percent")
rownames(volley) <- c("Current", "Former", "Never")
volley <- as.table(volley)
volley

```

**Cross Table**

```{r}
volley1 <- matrix(c(30, 20, 50, 60, 30, 90, 20, 20, 40), ncol = 3, byrow = TRUE)
colnames(volley1) <- c("Male", "Female", "Total")
rownames(volley1) <- c("Current", "Former", "Never")
volley1 <- as.table(volley1)
volley1
```

**From numeric to categorical**

```{r}
set.seed(123)
marital = sample(seq(from = 1, to = 2, by = 1), size = 150, replace = T) #data for marital status
membership = sample(seq(from = 1, to = 3, by = 1), size = 150, replace = T) #data for membership
data = data.frame(marital, membership) # Create a dataframe

glimpse(data)


```

```{r}
labs <- c("single", "married") # define the attributes for marital status
labs1 = c("former", "current", "never") # define attributes for membership
marital <- factor(data$marital, labels = labs) # assign attributes
membership = factor(data$member, labels = labs1) # assign attributes

marital;membership
```

```{r}
table <- xtabs(~membership + marital) 
table
```

```{r}
barplot(table, ylim=c(0,110), legend=T,main= 'Volleyball Membership',
args.legend = list(x="top"), col = c("skyblue", "red", "green"))

```

##Exercise

```{r}

library(stargazer) # Package made to export regression and summary tables
library(haven) # And a package for loading different data types fast(er)
#### Load some data that we want to test out
# Data comes from an experiment run by Bertrand and Mullainathan (2004)
# on the effect of having a non-black sounding name.
dat <- read_dta("http://masteringmetrics.com/wp-content/uploads/2015/02/bm.dta")
glimpse(dat)

datt= read_csv(file="C:/Users/Simon ik mig/Desktop/Statistik 2021/Statistik-3.semester/dat.csv" )
#More columns than column names. Drops the last column. 

glimpse(datt)
view(datt)
```

```{r}
labs1 = c("No interview call", "Interveiew call") # define attributes for interview call
labs2 = c("white american", "african american") # define attributes for race
interview = factor(dat$call, labels = labs1) # assign attributes
race = factor(dat$black, labels = labs2) # assign attributes
table <- xtabs(~interview + race)
table
```

```{r}
barplot(table, ylim=c(0,3200), legend=T,main= 'Race and call for job interviews',
args.legend = list(x="top"), col = c("skyblue", "yellow"))

```

**Pie Chart**

```{r}
ms <- matrix(c(37.69, 36.54, 16.03, 4.90, 4.26, 0.58), ncol = 1, byrow = T)
colnames(ms) <- c("EU market")
rownames(ms) <- c("Firefox","Google chrome", "Internet Explorer", "Safari", "Opera", "Others")
as.table(ms)
```

```{r}
pie(ms, labels = c("Firefox", "Google chrome", "Internet Explorer", "Safari", "Opera", "Others"))
```

```{r}
par(mar=c(2,10,10,10))
par(cex=1)
labels= c("Firefox", "Google chrome", "Internet Explorer", "Safari", "Opera", "Others")
pct <- round(ms/sum(ms)*100)
lbs= paste(labels, pct)
lbs= paste(lbs, "%",sep="")
pie3D(ms, labels = lbs, explode=0.1, radius=.8, theta = 1.2, labelcex = 1.4)
```

**Pareto Chart**

```{r}
counts <- matrix(c(12, 29, 18, 3, 34, 4), ncol = 1, byrow = T)
rownames(counts) <- c("Weather", "Overslept", "Alarm Failure", "Time Change",
"Traffic", "Other")
colnames(counts) <- c("Reason for being late")
counts
```

```{r}

pareto.chart(counts[,1])

```

#Graphs to descripe Numerical values

##Histograms

```{r}
x <- rnorm(1000, mean=0, sd = 1); y <- rf(1000, 10, 20); z <- -rf(1000, 10, 20)

hist(x, col = rgb(0.8, 0.8, 1), main = "Normal distribution")
hist(y, col = rgb(0.8, 0.8, 1), main = "Right-skewed")
hist(z, col = rgb(0.8, 0.8, 1), main = "Left-skewed")


```

```{r}
hist(x, breaks = 25, col = rgb(0.8, 0.8, 1), main = "Normal distribution")

```

##Scatterplot

```{r}
attach(mtcars) # this contains a dataset which is included in R by default

plot(wt, mpg, main="Scatterplot", xlab="Car Weight ", ylab="Miles Per Gallon ", pch=16)

```

```{r}
scatterplot3d(wt, disp, mpg, pch=16, highlight.3d = TRUE, type="h", main="3D Scatterplot")
```

##Time series plots

```{r}
library(Ecdat)
data(Money) # This data matrix contains time series of interest rates
view(Money)

interest = Money[, 4]

interest = ts(interest, start = c(1967, 1), end = c(1998, 4), frequency = 4)

plot(interest, col="blue", main="Interest rate")

```

#Exercise

##R-Markdown

MikTeX url: <https://miktex.org/download>

```{r}
#install.packages('tinytex')
#tinytex::install_tinytex()


```

##Lecture exercise

1.  make a contigency table and bar chart of **computer skills** and **gender** in the following data

```{r}
pc <- read_dta("http://masteringmetrics.com/wp-content/uploads/2015/02/bm.dta")

glimpse(pc)

pc_label= c("yes", "no")
pc_label2= c("male", "female")

computer_skills= factor(pc$computerskills, labels = pc_label)
gender= factor(pc$female, labels = pc_label2)

pc_table= xtabs(~computer_skills + gender)

pc_table

```

We can create a Bar chart:

```{r}
barplot(pc_table, ylim=c(0,3800), legend=T,main= 'Computer skills',
args.legend = list(x="left"), col = c("skyblue", "red"))
```



2. enerate 1000 random numbers in R (show data distribution). Comment on data distribution

```{r}
random_num=rnorm(1000, 0, 1) #We make them as a normal distribution 

hist(random_num, col = rgb(0.8, 0.8, 1), main = "Normal distribution")

num=runif(1000, min=0, max=100)

hist(num, col = rgb(0.8, 0.8, 1), main = "Uniform distribution") #more like a uniform

```

3. Create your own cross table in R (show a bar plot of your cross table)


Look in Starwars


4. Generate 100 random numbers. Format the data as annual time series from 1915 to 2015. Plot the annual time series.

```{r}
ts_data=rnorm(100, 0, 1)

ts_data=ts(ts_data, start = c(1915), end = c(2015), frequency = 1)

plot(ts_data, main="Random numbers" )

```

5. Generate 240 random numbers. Format the data as monthly time series from 1990M1 to 2010M12. Plot the monthly time series.



```{r}
ts_data1=rnorm(240, 0, 1)

ts_data1=ts(ts_data1, start = c(1990, 1), end = c(2010,12), frequency = 12)

plot(ts_data1, main="Random numbers")
```


























