---
title: "Lektion 2"
author: "Simon"
date: "22/9/2021"
output: html_document
---

```{r}
library(tidyverse)
```


# Numerical measures

**calculating percentiles and Quartiles**

```{r}
sales = c(85, 84, 80, 82, 60, 63, 65, 67, 70, 72, 75, 75)%>%sort()

```
```{r}
quantile(sales, type = 6)
```
Calculate 44% percentile
```{r}
quantile(sales, type = 6, prob = c(0.44))
```

**Box-and-Whisker Plot (Cont)**

```{r, fig.height=7, fig.width=10}
set.seed(123)

copenhagen = sample(seq(from = 10, to = 40, by = 3), size = 15, replace = T)
zealand = sample(seq(from = 5, to = 25, by = 2), size = 15, replace = T)
n_jutland = sample(seq(from = 5, to = 30, by = 3), size = 15, replace = T)
central_jutland = sample(seq(from = 5, to = 35, by = 4), size = 15, replace = T)
southern_denmark = sample(seq(from = 2, to = 28, by = 6), size = 15, replace = T)
data = data.frame(copenhagen, zealand, n_jutland, central_jutland, southern_denmark)
boxplot(data, lwd=1, main="Box-plot", xlab="Regions ", ylab="Sales",
names=c("Copenhangen","Zealand","Northern Jutland","Central Jutland","Southern Denmark"))

```
We can make it in colors to make it more fancy 


```{r, fig.height=7, fig.width=10}

boxplot(data, lwd=1, main="Box-plot", xlab="Regions ", ylab="Sales",
names=c("Copenhangen","Zealand","Northern Jutland","Central Jutland","Southern Denmark"),
col = c("red","sienna","palevioletred1","royalblue2","yellow"))

```

Task: 
- Calcculate 1st quartile, second quartile, and third quartile of the regions ploted 

```{r}
apply(data, 2,  quantile)


data$copenhagen%>% quantile(type= 6)
data$zealand%>% quantile(type= 6)
data$n_jutland %>% quantile(type= 6)
data$central_jutland%>% quantile(type= 6)
data$southern_denmark%>% quantile(type= 6)


```

## Exercise


*Assume two variables $x$ and $y$ in R:
```{r}
set.seed(213)
x = sample(seq(from = 10, to = 1000, by = 3), size = 1241, replace = T)
y = sample(seq(from = 10, to = 100, by = 3), size = 1241, replace = T)

```


*What is the number of observations?

```{r}
n_x=length(x); n_y=length(y)

```

*Show boxer and whisker plot of the two series. What do you see?

```{r}
data_xy=cbind(x,y)

boxplot(data_xy, lwd=1, main="Box-plot", xlab="data ", ylab="numbers",
names=c("x","y"),
col = c("red","sienna"))


```

as it looks very bad we make one for each:

```{r}
boxplot(x, lwd=1, main="Box-plot", xlab="data ", ylab="numbers",
names=c("x"),
col = c("red"))

boxplot(y, lwd=1, main="Box-plot", xlab="data ", ylab="numbers",
names=c("y"),
col = c("red"))
```

*What is the mean of $x$ and $y$ compare your results with mean(x) in R?

```{r}

mean_x=sum(x)/n_x; mean(x)

mean_y=sum(y)/n_y; mean(y)

```

*Calculate the variance of $x$ and $y$, compare your results with var(x) in R:

```{r}

var_x=(sum((x-mean_x)^2))/(n_x-1);var_x

var_y=(sum((y-mean_y)^2))/(n_y-1);var_y


```
*Calculate standard deviation of x and y, compare your results with sd(x) in R:

```{r}
sd_x=sqrt(var_x)

sd_y= sqrt(var_y)

```
*Finally, calculate coefficient of variance for x and y.

```{r}
cv_x= sd_x/mean_x;cv_x

cv_y= sd_y/mean_y;cv_y
```
*Assume x and y were return on assets. Which asset would you invest in?

The one with lowest CV as this is the standard deviation compared with the mean. 


# Basic Probability

