---
title: "Lektion 3"
author: "Simon"
date: "30/9/2021"
output: html_document
---


# Random Variable 



# Probability distribution function of discrete random variables

```{r}
px = c(0.1, 0.2, 0.4, 0.3); barplot(px, names.arg = c(0,1,2,3))
```


# Properties of discrete random variables

**Expected value of a discrete Random Variable**

```{r}
x = c(0, 1, 2, 3, 4, 5); px = c(0.15, 0.3, 0.2, 0.2, 0.1, 0.05)

mean = sum(x * px)
mean

```
**Variance of a discrete random variable**

```{r}
sum((x - mean)^2 * px)
```

##Example Expected value of functions of random variables. 

```{r}
x = c(10, 11, 12, 13, 14); px = c(0.1, 0.3, 0.3, 0.2, 0.1); c = 25000 + 900*x
```

We can easily calculate the mean of X:

```{r}
mean_x = sum(x * px)

```

We can also easily calculate the mean of total cost C; $µy = E[a + bX] = a + bµx$

```{r}
mean_c = 25000 + 900 * mean_x
```

We can easily calculate the variance of total cost C $\sigma^2_y= Var[a+bX]= b^2\sigma ^2_x$

```{r}
var_x = sum((x - mean_x)^2*px) # first calculate the variance of X
var_c = (900^2)*(var_x)
var_c
```

# Binomial Distribution

To compute the binomial probabilities in R, we use dbinom. The probability of getting 1
head in two coin tosses is:


```{r}
dbinom(1, size=2, prob=0.5)

#The probabilities of getting 0, 1, 2 heads in two trials is:

dbinom(0:2, size=2, prob=0.5)
```
**Number of sequences with x Successes in n Trails**

```{r}
choose(5,2) #In 10 sequences we can hit 2 heads out of 5 trails. 
```
##Example

Suppose that a property dealer, John, has 5 contacts, and he believes that for each
contact the probability of making a sale is 0.40.


**a. Find the probability that he makes 1 sale**

```{r}
dbinom(1, size = 5, prob = 0.4)
```


**b. What is the probability that he makes at most 1 sale**

```{r}
sum(dbinom(0:1, size = 5, prob = 0.4))
```


**c. Find the probability that he makes between 2 and 4 sales (inclusive).**

```{r}
p2 = dbinom(2, size = 5, prob = 0.4)
p3 = dbinom(3, size = 5, prob = 0.4)
p4 = dbinom(4, size = 5, prob = 0.4)
p2 + p3 + p4

sum(dbinom(2:4, size = 5, prob = 0.4))
```


**d. Graph the probability distribution function.**

```{r}
x=dbinom(0:5, size = 5, prob = 0.4)
```
```{r}
barplot(x, names.arg = c(0,1,2,3,4,5))

```
Lets try a few plots in R (note the probabilities):

```{r}
x <- dbinom(0:20, size = 20, prob = 0.2)
y <- dbinom(0:20, size = 20, prob = 0.8)
z <- dbinom(0:20, size = 20, prob = 0.5)
```

```{r}
barplot(x, col = rgb(0.8, 0.8, 1), main = "asymmetric")
barplot(y, col = rgb(0.8, 0.8, 1), main = "asymmetric")
barplot(z, col = rgb(0.8, 0.8, 1), main = "symmetric")
```

**Cumulative Binomial Distribution in R**

```{r}
x <- dbinom(0:5, size = 5, prob = 0.4)
x
```
```{r}
x <- pbinom(0:5, size = 5, prob = 0.4)
x
```

**Mean and Variance of binomial distribution**

```{r}
x <- 0:2
px = dbinom(x, size = 2, prob = 0.5)
mean=sum(x * px)

var=5*0.4*0.6


```


# Exercise

**Calculate the probability of 13 heads in 30 independent bernoulli trials**

```{r}
x=13
dbinom(13, size = 30, prob = 0.5)

```


**Calculate the binomial distribution of the above experiment**

```{r}
y= dbinom(0:30, size = 30, prob = 0.5)
barplot(y, names.arg = c(0:30))
```


**Calculate the cummulative distribution of the above experiment**

```{r}
y= dbinom(0:30, size = 30, prob = 0.5)
z= cumsum(y)
```

```{r}
par(mfrow=c(1,2))
barplot(y, names.arg = c(0:30))
barplot(z, names.arg = c(0:30))
```

**Calculate the mean and variance of heads in 30 bernoulli trials**
Mean:
```{r}
x=0:30
px= dbinom(x, size = 30, prob = 0.5)
sum(x*px)
```
Variance:

```{r}
mean_x <- sum(x*px)
sum((x-mean_x)^2*px) #Methode from discrete variables

30*0.5*0.5 #Can also use the rule when its binomial.
```




