---
title: "Lektion 2"
author: "Simon"
date: "22/9/2021"
output: html_document
---

```{r}
library(tidyverse)
```


# Numerical measures

**calculating percentiles and Quartiles**

```{r}
sales = c(85, 84, 80, 82, 60, 63, 65, 67, 70, 72, 75, 75)%>%sort()

```
```{r}
quantile(sales, type = 6)
```
Calculate 44% percentile
```{r}
quantile(sales, type = 6, prob = c(0.44))
```

**Box-and-Whisker Plot (Cont)**

```{r, fig.height=7, fig.width=10}
set.seed(123)

copenhagen = sample(seq(from = 10, to = 40, by = 3), size = 15, replace = T)
zealand = sample(seq(from = 5, to = 25, by = 2), size = 15, replace = T)
n_jutland = sample(seq(from = 5, to = 30, by = 3), size = 15, replace = T)
central_jutland = sample(seq(from = 5, to = 35, by = 4), size = 15, replace = T)
southern_denmark = sample(seq(from = 2, to = 28, by = 6), size = 15, replace = T)
data = data.frame(copenhagen, zealand, n_jutland, central_jutland, southern_denmark)
boxplot(data, lwd=1, main="Box-plot", xlab="Regions ", ylab="Sales",
names=c("Copenhangen","Zealand","Northern Jutland","Central Jutland","Southern Denmark"))

```
We can make it in colors to make it more fancy 


```{r, fig.height=7, fig.width=10}

boxplot(data, lwd=1, main="Box-plot", xlab="Regions ", ylab="Sales",
names=c("Copenhangen","Zealand","Northern Jutland","Central Jutland","Southern Denmark"),
col = c("red","sienna","palevioletred1","royalblue2","yellow"))

```

Task: 
- Calcculate 1st quartile, second quartile, and third quartile of the regions ploted 

```{r}
apply(data, 2,  quantile)


data$copenhagen%>% quantile(type= 6)
data$zealand%>% quantile(type= 6)
data$n_jutland %>% quantile(type= 6)
data$central_jutland%>% quantile(type= 6)
data$southern_denmark%>% quantile(type= 6)


```

## Exercise


*Assume two variables $x$ and $y$ in R:
```{r}
set.seed(213)
x = sample(seq(from = 10, to = 1000, by = 3), size = 1241, replace = T)
y = sample(seq(from = 10, to = 100, by = 3), size = 1241, replace = T)

```


*What is the number of observations?

```{r}
n_x=length(x); n_y=length(y)

```

*Show boxer and whisker plot of the two series. What do you see?

```{r}
data_xy=cbind(x,y)

boxplot(data_xy, lwd=1, main="Box-plot", xlab="data ", ylab="numbers",
names=c("x","y"),
col = c("red","sienna"))


```

as it looks very bad we make one for each:

```{r}
boxplot(x, lwd=1, main="Box-plot", xlab="data ", ylab="numbers",
names=c("x"),
col = c("red"))

boxplot(y, lwd=1, main="Box-plot", xlab="data ", ylab="numbers",
names=c("y"),
col = c("red"))
```

*What is the mean of $x$ and $y$ compare your results with mean(x) in R?

```{r}

mean_x=sum(x)/n_x; mean(x)

mean_y=sum(y)/n_y; mean(y)

```

*Calculate the variance of $x$ and $y$, compare your results with var(x) in R:

```{r}

var_x=(sum((x-mean_x)^2))/(n_x-1);var_x

var_y=(sum((y-mean_y)^2))/(n_y-1);var_y


```
*Calculate standard deviation of x and y, compare your results with sd(x) in R:

```{r}
sd_x=sqrt(var_x)

sd_y= sqrt(var_y)

```
*Finally, calculate coefficient of variance for x and y.

```{r}
cv_x= sd_x/mean_x;cv_x

cv_y= sd_y/mean_y;cv_y
```
*Assume x and y were return on assets. Which asset would you invest in?

The one with lowest CV as this is the standard deviation compared with the mean. 


# Basic Probability


**Union of events**

```{r}
A=c(1,3,5); B=c(2,4,6)
union(A,B)
```
**Intersections of events**

```{r}

A=c(1,2,3,5,6); B=c(2,4,6,8,9)
intersect(A,B)

```

**probability Rules (cont)**

Example 1: 50 people were surveyed to answer "What are the top 2 reasons you
are late to work?". The responses were following:

```{r}
counts <- matrix(c(6, 14, 9, 2, 17, 2), ncol = 1, byrow = T)
rownames(counts) <- c("Weather", "Overslept", "Alarm Failure", "Time Change", "Traffic", "Other")
colnames(counts) <- c("Reason for being late")
counts

```
We can calculate the relative frequency of the above example:

```{r}
100 * prop.table(counts, 2)

```
Example 2: For example, employees of a company are asked their membership
status of a volleyball club in Aalborg as well as their gender:


```{r}
volley1 <- matrix(c(30, 20, 60, 30, 20, 20), ncol = 2, byrow = TRUE)
colnames(volley1) <- c("Male", "Female")
rownames(volley1) <- c("Current", "Former", "Never")
volley1 <- as.table(volley1)
100 * prop.table(volley1, 2)
```

```{r}
100 * prop.table(volley1, 1) # This calculates row-wise relative frequencies
```
```{r}
100 * prop.table(volley1) #Now its not column or row wise, so all 6 probabilities equals to 1 now


```

##Eercise

```{r}
library(tidyverse) # Data wrangling and plotting tools
library(stargazer) # Package made to export regression and summary tables
library(haven) # And a package for loading different data types fast(er)
#### Load some data that we want to test out Data comes from an experiment
#### run by Bertrand and Mullainathan (2004) on the effect of having a
#### non-black sounding name.
dat <- read_dta("http://masteringmetrics.com/wp-content/uploads/2015/02/bm.dta")
```



**Calculate a contigency table between education and interview call**

```{r}
labs1 = c("not", "primary", "high", "college", "university") # define attributes for interview call
labs2 = c("no call", "call") # define attributes for race
education = factor(dat$education, labels = labs1) # assign attributes
interview = factor(dat$call, labels = labs2) # assign attributes
table <- xtabs(~education + interview)
table
```

**Calculate row wise probabilities**

```{r}
100 * prop.table(table, 1)
```
**Calculate column wise probabilities**


```{r}
100 * prop.table(table, 2)
```
**Calculate discrete distribution.**

```{r}
100 * prop.table(table)
```


## Conditional probability (cont) Exercise

A company gathers data on the gender of its employees and their volleyball membership status

```{r}
d_prob= prop.table(volley1)
```

**What is the probability that an employee is a male or current member of the volleyball club, i.e., find P(male ∪ current)?**

```{r}

p_male_current= d_prob[1,1]
p_male= sum(d_prob[,1])
p_current= sum(d_prob[1,])

p_male + p_current - p_male_current

```
**What is the probability that an employee is a female or a former member of the volleyball club, i.e., find P(female ∪ former)?**

```{r}

p_female_former= d_prob[2,2]
p_female= sum(d_prob[,2])
p_former= sum(d_prob[2,])

p_female + p_former - p_female_former

```

**What is the probability that an employee who is a male is also a former club member?**

```{r}
d_prob[2,1]/p_male
```
**What is the probability that an employee who is a female is also a current club member?**

```{r}
p_female= sum(d_prob[,2])

d_prob[1,2]/p_female 
```


# Bayes Theorem

## Exercises

**Given P(A1) = 0.4, P(B1|A1) = 0.6, and P(B1|A2) = 0.7, what is the probability of P(A1|B1)?**

We need to use the equation from slide 32

```{r}
p_a1 = 0.4
p_b1_a1 = 0.6
p_b1_a2 = 0.7
p_a2= 1-p_a1

(p_b1_a1*p_a1)/(p_b1_a1*p_a1+p_b1_a2*p_a2)


```

**Given P(A1) = 0.4, P(B1|A1) = 0.6, and P(B1|A2) = 0.7, what is the probability of P(A2|B2)?**

We need the same function as we do not know b2

```{r}
p_b2_a2= 1- p_b1_a2
p_b2_a1= 1-p_b1_a1
p_a2= 1-p_a1

(p_b2_a2*p_a2)/ (p_b2_a2*p_a2 + p_b2_a1*p_a1)

```


**A publisher sends advertising materials for an accounting text to 80% of all professors teaching the appropriate accounting course. 30% of the professors who received this material adopted the book, as did 10% of the professors who did not receive the material. What is the probability that a professor who adopts the book has received the advertising material?**

```{r}

#R= recived the book
# A= adopted the book

r1= 0.8
r2= 1- r1

a1_r1= 0.3
a2_r1 = 1-a1_r1

a1_r2 = 0.1
a2_r2 = 1-a1_r2

#We can now find r1_a1
#Since we do not have p(a1) we need to use thhe same function as before

r1_a1= (a1_r1*r1)/(a1_r1*r1+a1_r2*r2)
```



