---
title: "Lektion 2 Example"
author: "Simon"
date: "23/9/2021"
output: html_document
---


```{r}
library(tidyverse) 
library(magrittr)
library(data.table)
library(skimr)
```


```{r}
data <- read_csv("https://sds-aau.github.io/SDS-master/M1/data/RI-clean.csv.gz")
```

```{r}
data %>% head()
```

```{r}
data %>% glimpse()
```

```{r}
data%<>% 
  select(driver_gender, violation)
```


Drop na's

```{r}
data%<>% drop_na()

data %<>%
  group_by(violation) %>%
  count(driver_gender)%>%
  mutate(violation=as.factor(violation), driver_gender= as.factor(driver_gender))

data <- xtabs(n ~ violation + driver_gender, data = data)




```




## Conditional probability (cont) Exercise

A company gathers data on the gender of its employees and their volleyball membership status

```{r}
d_prob=100 * prop.table(data)
```

**What is the probability that an employee is a male or current member of the volleyball club, i.e., find P(male ∪ current)?**

```{r}
p_male_seatbelt= d_prob[1,1]
p_male= sum(d_prob[,2])
p_seatbelt= sum(d_prob[5,])

p_male + p_seatbelt - p_male_seatbelt 

```
**What is the probability that an employee is a female or a former member of the volleyball club, i.e., find P(female ∪ former)?**



**What is the probability that an employee who is a male is also a former club member?**

```{r}
d_prob[2,1]/p_male
```
**What is the probability that an employee who is a female is also a current club member?**

```{r}
p_female= sum(d_prob[,2])/100

d_prob[1,2]/p_female 
```

